### 2022-04-09 ret_oriented_prog.tex
added the basic template

\documentclass[10pt,a4paper,twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\author{Maximilian Heim}
\title{Exploring Return Oriented Programming}
\begin{document}
\maketitle
\section{Abstract}
\section{Introduction}
\section{Conclusion}
\bibliography{bib.bib}
\end{document}